ShrinkNets

### 2018-01-05 paper/paper.tex
pass

When designing Neural Networks, finding the appropriate size
%, number of layers
%and number of neuros per layer, 
%(width and depth)
is key. It consists of determining the number of layers and neurons per layer
that lead to the best accuracy.
%In particular, these hyper parameters have a strong correlation with
%over/underfitting \srm{give reference} \gl{I did some litterature review and
%actually this statement is quite controversial, should we rephrase ?}. 
Unfortunately, we have no reliable way to find the best parameters. Decades of
experimentation led to search strategies and heuristics~\cite{Bengio2012a} that prune
the immense space of possible network sizes and suggest range of values to
explore. Examples of search strategies are
%To select the best candidates, researchers have used strategies such
%as 
random search,
%\gl{Should we also move the references here if we remove the
%pargraph from the related work ?}, 
meta-gradient descent~\cite{Pedregosa2016},
gaussian processes~\cite{Bergstra2011a}, and Parzen Estimators~\cite{Bergstra2011a}. 
%to find good parameters without exhaustively
%exploring the entire hyper-parameter space. 
Although these strategies help with
finding a good set of hyper-parameters, they still require a compute-intensive
search of the space.

@@

In this paper we present a method to automatically find an appropriate network
size, thus
%from a single parameter, 
drastically reducing the hyper-parameter optimization time.
%dimensionality. 
The key idea is to
\emph{learn} the right network size at the same time that the network is
learning the main task. For example, for an image classification task, with our
approach we can provide the training data to a network---without sizing it a
priori---and expect to end up with a network that has learnt to classify images
with an accuracy similar to a manually engineered network.
%\gl{too strong, we do not really have
%  any guarantee on the generalization, how about changing it to: to end up with
%  a network that learnt a tradeoff between size and accuracy}. 
This approach has two main benefits. First, we do no longer need to choose a
network size before training. Second, the final network size will be
appropriate for the task at hand, and not larger. 
%Because the sizing occurs
%dynamically, while the
%network is training, we remove the useless neurons 
%implement a new mechanism that removes
%the deactivated neurons to  
%\gl{Should we talk about the
%fact that it automatically gets rid of dead neurons ?}. 
This is important
because oversized networks have a lower inference throughput and higher memory
footprint.

Our approach has two main challenges. First, on how to dynamically size the
network during training. Second, on how to find a loss function that optimizes
for the additional task of sizing, without deteriorating the learning
performance of the main task. We describe next ShrinkNets, which cope with both
challenges:

@@

Our approach consists of starting the training process for the task of interest
with an explicitly oversized network. Then, as training progresses, we learn
which neurons are not contributing to learning the main task and remove them
dynamically, thus shrinking the network size. This method requires two building
blocks. First, a way of identifying neurons that are not contributing to the
learning process, and second a way of balancing the network size and the
generalization capability for the main task. We introduce a new layer, called
\textsf{Filter}, which takes care of \emph{deactivating} neurons. We also modify
existing loss functions to incorporate a new term that takes care of balancing
network sizing and generalization capability appropriately. We explain them
next:

@@

\textbf{Filter Layers: } They have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers of traditional deep networks.
%\gl{This is how I
%  used it but with the new implementation we can be more creative and put them
%  anywhere, should we say "usually placed" insdead ?}. 
The \textit{Filter Layer} takes an input of size $\left(B \times C \times D_1
\times \dots \times D_n\right)$, where $B$ is the batch size, $C$ the number of
features (or channels, in the case of convolutional layers), and $D$ any additional dimension. This structure makes
it compatible with fully connected layers with $n=0$ or convolutional layers
with $n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$,
defined as follows:

\begin{equation} Filter(I;\theta) = I \circ \max(0, \theta) \end{equation}

Where $\circ$ is the 
%Hadamard product (
pointwise multiplication,
%\gl{Should we
%  keep only one of them to be more concise ?}), 
and $\theta$ is expanded in all
dimensions except the second one to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer deactivates the
neuron. These disabled neurons/channels can be removed from the network without
changing its output. Before explaining how that is achieved, we 
% (we describe how we perform this removal below). We 
explain
next how the weights of the Filter Layer are adjusted during training.

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a deep network, 
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train it directly and it would be equivalent to a normal neural
network. However, our goal is to find the smallest network with reasonable
performance. We achieve that by introducing sparsity in the parameters of the
\textit{Filter Layers}, thus forcing the deactivation of neurons.
%. Indeed, having a negative component in the $\theta$
%parameter of the filter layer permamently disable its associated feature
%\gl{Maybe redundant ? we talked about that in the previous paragraph} . 
To obtain this sparsity, we simply redefine the loss function:

\begin{equation}
  L'(x,y;\theta) = L(x, y) + \lambda|\theta|
\end{equation}

The additional term $\lambda|\theta|$ introduces sparsity (see Lasso
loss~\cite{Tibshirani1996}).
%This choice comes from the Lasso loss ~\cite{Tibshirani1996} and we use it for
%the same reason: it introduces sparsity. 
The $\lambda$ parameter adjusts how aggressive the network deactivates neurons,
with larger values indicating more aggressive deactivation. 
%The bigger the $\lambda$ the more
%entries in $\theta$ are set to zero, and 
Because the entries in $\theta$ with a value of $0$
correspond to dead neurons, $\lambda$ effectively controls the number of
neurons/channels in the entire network.  
Next, we explain how to implement
ShrinkNets efficiently.

\textbf{Dynamic Network Resizing}
It is possible to reduce the overhead of the training process by removing
neurons as soon as they become deactivated.
%If
%disabled neurons are not quickly removed, the overhead might cause the training
%process to be significantly slower than classic neural networks. 
For that, we implemented a \emph{neural garbage collection} mechanism which
prunes deactivated neurons on-the-fly, therefore reducing the processing time
and memory overhead. To support this feature, it is crucial to understand the
information flow between neurons and layers in the neural network. We achieve
this by representing such information flow as a graph, in which edges indicate
what action each node, i.e., neuron, must take.

@@

We have implemented ShrinkNets, including the \emph{garbage collection}
mechanism on top of \textit{PyTorch}~\cite{paszke2017automatic}, and we have
made the software open source.\footnote{The code is available as Python/PyTorch
library on \url{http://github.com/mitdbg/fastdeepnets} 
%\gl{Should we rename the repository ?}
}. We have then used our implementation to evaluate two crucial aspects of
ShrinkNets, which we present next.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.9\columnwidth]{convergence}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\end{center}
\end{figure}

\subsection{Relevance in the context of hyper-parameter optimization}

One of our main goal is to help finding neural network architectures that
perform reasonably well faster than existing techniques. For simplicity, we
will try to see how ShrinkNets perform when doing random search over a sub set
of the hyper-parameter space (we will fix everything except the size-related
parameters). However, we believe that the results generalize to more complex
methods, most of them might actually benefit a lot from the reduction of the
dimensionality of the hyper-parameter space.

\par To isolate the effect of ShrinkNets we will consider two very simple
architectures: multi-layer perceptrons (with three hidden layers) and the
\textit{LeNet-5} model ~\cite{Lecun1998}. For this experiment we assume we have no prior
information about the size except an upper bound of $50$ channels per
convolutional layer and $5000$ neuron for fully connected layers.  We try to
find the appropriate size of the network using classical neural networks and
ShrinkNets. To make sure that we are as fair as possible and since one might
argue that ShrinkNets is also regularizing the network in addition to finding
the size, we will also considered regularized classic networks using the most
commonly used technique: the $L_2$ penalty ~\cite{Ng2004}.
\par The experiment consist in sampling parameters randomly for each class of
network, train them for the same amount of epochs (100 for MLPs and 200 for CNNs). Pick the epoch that performed the best on the validation set and evaluate it
on the testing set. We repeat the process for 50 models. This evaluation was done
on \texttt{MNIST} ~\cite{Lecun1998}, \texttt{FashionMNIST} ~\cite{Xiao2017} and \texttt{CIFAR10} ~\cite{Krizhevsky2009}. The distribution of the testing accuracy we obtained is summarized on \autoref{hyper_opt_res}.

We can see that the distributions for ShrinkNets are consistently better than the
others for all datasets and both architectures. It shows that by training fewer
networks, we are likely to obtain on-par or even better models which effectively
reduces the cost of hyper-parameter optimization. In some cases the best ShrinkNets
models even outperforms the best model of the other techniques.

\begin{figure}
\begin{center}
  \includegraphics[width=0.9\columnwidth]{hyper_opt}
\caption{Distributions of the testing accuracy for different training methods, datasets and architectures using random search\label{hyper_opt_res}}
\end{center}
\end{figure}

@@

\par There are many methods in the litterature that try to simplify the neural
network structure. Most of them focuses on removing connections (eg: ~\cite{Cun},
~\cite{Han2015}).  However, most hardware (GPU's, TPUS ~\cite{Jouppi2017}, TensorCores) do not really
leverage sparsity in the connections efficiently (as explained in ~\cite{Han2016}). On the other side, ShrinkNets
and some others ~\cite{Scardapane2017} and ~\cite{Philipp} try to remove entire neuron instead of
connections. This is very usefull because it reduces the size of the matrices,
producing speed-up on every device. ShrinkNets improves on
~\cite{Scardapane2017} because it removes neurons during training, speeding up
the rest of the process and beats ~\cite{Philipp} on convergence speed and because it
does not need to change the optimizer. To the best of our knowledge this is
also the first contribution that tries to learn the number of channels of a
convolutional neural network.

@@

In this paper we presented a novel technique to guess a resonable network size
based on single parameter $\lambda$ that control the tradeof between loss and
the size. We demonstrated that it works both on fully connected networks and
convolutional neural networks.  Even though the firsts results seem promising,
there are many ways we could improve. In the current implementation we only
"learn" the number of features (neurons or channels). We could try to augment
it with dynamic number of layers as seen in ~\cite{meier} to be able to
determine the entire architecture. \par We saw on Figure \ref{convergence_plot}
that the loss temporarily suffers from the removal of neurons. It is likely
that the loss would be more stable if the number of neurons converged faster or
neurons disappeared slower. For this reason we plan to explore proximal
gradient methods to optimize the filter vectors and/or randomize neuron
removals. \par During our evaluation we picked small datasets mainly to be able
to train many models and have statistically significant distributions. With
more computation resources and time, we could see if it generalizes to bigger
datasets and other architectures like \texttt{ResNet} ~\cite{He2016} (small
modifications to the existing code base are required to support them)

>>>>

When designing neural networks, one of the key parameters is the network size, i.e., 
%, number of layers
%and number of neuros per layer, 
%(width and depth)
the number of layers and neurons per layer.  Choosing these parameters appropriately
can dramatically affect performance, yet
%In particular, these hyper parameters have a strong correlation with
%over/underfitting \srm{give reference} \gl{I did some litterature review and
%actually this statement is quite controversial, should we rephrase ?}. 
there is no reliable way to efficiently set them. Although many search
 strategies and heuristics~\cite{Bengio2012a} have been proposed, including
%To select the best candidates, researchers have used strategies such
%as 
random search,
%\gl{Should we also move the references here if we remove the
%pargraph from the related work ?}, 
meta-gradient descent~\cite{Pedregosa2016},
Gaussian processes~\cite{Bergstra2011a}, and Parzen Estimators~\cite{Bergstra2011a}. 
%to find good parameters without exhaustively
%exploring the entire hyper-parameter space. 
 these methods generally require a compute-intensive
search of parameter space.

@@

In this paper we present a method to automatically find an appropriate network
size,
%from a single parameter, 
drastically reducing optimization time.
%dimensionality. 
The key idea is to
\emph{learn} the right network size at the same time that the network is
learning the main task. For example, for an image classification task, with our
approach we can provide the training data to a network---without sizing it a
priori---and expect to end up with a network that has learned to classify images
with an accuracy similar to a the best manually engineered network.
%\gl{too strong, we do not really have
%  any guarantee on the generalization, how about changing it to: to end up with
%  a network that learnt a tradeoff between size and accuracy}. 
Our approach has two main benefits. First, no longer need to choose a
network size before training. Second, the final network size will tuned to be
appropriate for the task at hand, and not larger. 
%Because the sizing occurs
%dynamically, while the
%network is training, we remove the useless neurons 
%implement a new mechanism that removes
%the deactivated neurons to  
%\gl{Should we talk about the
%fact that it automatically gets rid of dead neurons ?}. 
This is important
because over-sized networks have a lower inference throughput and higher memory
footprint.

Our approach has two main challenges. First, we need a way to dynamically size the
network during training. Second, we need a loss function that optimizes
for the additional task of sizing, without deteriorating the learning
performance of the main task. Our approach, called ShrinkNets, copes with both
challenges.

@@

During training, our approach starts
with an explicitly over-sized network. As training progresses, we learn
which neurons are not contributing to learning and remove them
dynamically, effectively shrinking the network. This method requires two key components: first, we need a way to identify neurons that are not contributing to the
learning process, and second we need a way to balance the network size and the
generalization capability for the main task. We introduce a new \textsf{Filter} layer that takes care of \emph{deactivating} neurons. We also modify
existing loss functions to incorporate a new term that takes care of balancing
network size and generalization capability appropriately.

@@

\textbf{Filter Layers: } Filter layers have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers.
%\gl{This is how I
%  used it but with the new implementation we can be more creative and put them
%  anywhere, should we say "usually placed" insdead ?}. 
The \textit{Filter Layer} takes an input of size $\left(B \times C \times D_1
\times \dots \times D_n\right)$, where $B$ is the batch size, $C$ the number of
features (or channels, in the case of convolutional layers), and $D$ any additional dimension. This structure makes
it compatible with fully connected layers with $n=0$ or convolutional layers
with $n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$,
defined as follows:
\vspace{-1em}
\begin{equation} Filter(I;\theta) = I \circ \max(0, \theta) \end{equation}
Where $\circ$ is the 
%Hadamard product (
pointwise multiplication,
%\gl{Should we
%  keep only one of them to be more concise ?}), 
and $\theta$ is expanded in all
dimensions except the second one \srm{why not the second one? what is the second one?  C?} to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer deactivates the
neuron. \srm{Not easy to see this because you haven't told me that theta only decays over time?} These disabled neurons/channels can be removed from the network without
changing its output. Before explaining how that is achieved, we 
% (we describe how we perform this removal below). We 
explain
next how the weights of the Filter Layer are adjusted during training. \srm{How are filter weights initialized?}

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a network, 
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train the network directly using our standard loss function, and we could achieve performance equivalent to a normal neural
network. However, our goal is to find the smallest network with reasonable
performance. We achieve that by introducing sparsity in the parameters of the
\textit{Filter Layers}, thus forcing the deactivation of neurons \srm{by decaying the value of theta}
%. Indeed, having a negative component in the $\theta$
%parameter of the filter layer permamently disable its associated feature
%\gl{Maybe redundant ? we talked about that in the previous paragraph} . 
To obtain this sparsity, we simply redefine the loss function:
\vspace{-.5em}
\begin{equation}
  L'(x,y;\theta) = L(x, y) + \lambda|\theta|
\end{equation}

The additional term $\lambda|\theta|$ introduces sparsity (see Lasso
loss~\cite{Tibshirani1996}). \srm{Say lambda is in [0,1]?}
%This choice comes from the Lasso loss ~\cite{Tibshirani1996} and we use it for
%the same reason: it introduces sparsity. 
The $\lambda$ parameter adjusts how aggressively the network deactivates neurons,
with larger values indicating more aggressive deactivation. \srm{Give some intuitive statement about how this loss function causes theta to decay in irrelevant neurons, while preserving the value of important neurons.}
%The bigger the $\lambda$ the more
%entries in $\theta$ are set to zero, and 
Because the entries in $\theta$ with a value of $0$
correspond to dead neurons, $\lambda$ effectively controls the number of
neurons/channels in the entire network.  
Next, we explain how to implement
ShrinkNets efficiently.

\textbf{Dynamic Network Resizing}
It is possible to reduce the overhead of the training process by removing
neurons as soon as they become deactivated by $\theta$ going to 0.
%If
%disabled neurons are not quickly removed, the overhead might cause the training
%process to be significantly slower than classic neural networks. 
To do this, we implemented a \emph{neural garbage collection} mechanism which
prunes deactivated neurons on-the-fly,  reducing the processing time
and memory overhead. To support this feature, it is crucial to understand the
information flow between neurons and layers in the neural network. We achieve
this by representing such information flow as a graph, in which edges indicate
what action each node, i.e., neuron, must take. 
\srm{I can't really understand what this means -- need to add a bit more detail.}

@@

We  implemented ShrinkNets, including the \emph{garbage collection}
mechanism on top of \textit{PyTorch}~\cite{paszke2017automatic}, and we have
made the software open source.\footnote{The code is available as Python/PyTorch
library on \url{http://github.com/mitdbg/fastdeepnets} 
%\gl{Should we rename the repository ?}
}. We used used our implementation to evaluate two crucial aspects of
ShrinkNets, which we present next.

@@

\begin{figure}
\vspace{-.2in}
\begin{center}
\includegraphics[width=0.9\columnwidth]{convergence}
\vspace{-.2in}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\srm{Move legend inside top of figure, to save space.}
\vspace{-.2in}
\end{center}
\end{figure}

\subsection{Does ShrinkNet find good networks?}

One of our main goals is to show that ShrinkNets quickly finds neural network architectures that
perform well. To show this, we ran a simple experiment where we compare ShrinkNets
to other network tuning methods,
 using random search over the
hyper-parameters that control the network size ($\lambda$ in the case of ShrinkNets). 
Despite the simple nature of this experiment, 
we believe that the results generalize to more complex
search methods, many of of which will actually benefit
dimensionality of the hyper-parameter space in the ShrinkNets approach.

To isolate the effect of ShrinkNets we will consider two very simple
architectures: multi-layer perceptrons (with three hidden layers) and the
\textit{LeNet-5} model ~\cite{Lecun1998}.
For this experiment we assume we have no prior
information about the size except an upper bound of $50$ channels per
convolutional layer and $5000$ neurons for fully connected layers.  We try to
find the appropriate size of the network using classical network-size optimization methods
 and
ShrinkNets. For the classical methods we use: i) purely random selection of network size parameters (``Fixed''), as well as hyper-parameter
regularization over the network size parameters using  $L_2$ penalty ~\cite{Ng2004} 
(``Fixed-L2'').
In the experiment, we randomly sample parameters for each class of
network and  train them for the same number of epochs (100 for MLPs and 200 for CNNs). We picked the epoch that performed the best on the validation set and evaluated it
on the testing set. We repeat the process for 50 iterations. This evaluation was done
on \texttt{MNIST} ~\cite{Lecun1998}, \texttt{FashionMNIST} ~\cite{Xiao2017} and \texttt{CIFAR10} ~\cite{Krizhevsky2009}. The distribution of the testing accuracy we obtained is summarized in \autoref{hyper_opt_res}.

We can see that in all cases the distributions for ShrinkNets have substantially higher medians are are less variable than the other architectures (the best model is not always found by ShrinkNets, but this is likely an artifact of our randomized search process).
This shows that ShrinkNets can find better networks with fewer training iterations than existing methods. \srm{Note something about how we expect this to bear out even with methods other than random search?}

\begin{figure}
\vspace{-.2in}
\begin{center}
  \includegraphics[width=0.9\columnwidth]{hyper_opt}
  \vspace{-.2in}
\caption{Distributions of the testing accuracy for different training methods, datasets and architectures using random search\label{hyper_opt_res}}
\vspace{-.2in}
\end{center}
\end{figure}

@@

\par There are many methods in the literature that aim to simplify 
network structure. Most of them focuses on removing connections (eg: ~\cite{Cun},
~\cite{Han2015}).  However, most hardware (GPU's, TPUS ~\cite{Jouppi2017}, TensorCores) do not really
leverage sparsity in the connections efficiently (as explained in ~\cite{Han2016}). On the other side, ShrinkNets
and some others ~\cite{Scardapane2017} and ~\cite{Philipp} try to remove entire neurons instead of
connections. This is very useful because it reduces the size of the matrices,
producing speed-up on every device. ShrinkNets improves on
~\cite{Scardapane2017} because it removes neurons during training, speeding up
the rest of the process.  Further it outperforms~\cite{Philipp} on convergence speed and because it does not need to change the optimizer. To the best of our knowledge this is
also the first work that tries to learn the number of channels in a
convolutional neural network.

@@

In this paper we presented a novel technique to guess a reasonable network size
based on single parameter $\lambda$ that control the tradeoff between loss and
network size. We demonstrated that ShrinkNets works both on fully connected networks and
convolutional neural networks.  Although the initial results are promising,
there are many additional avenues to explore. In the current implementation we only
``learn'' the number of features (neurons or channels). We plan to augment
this with a dynamic number of layers as seen in ~\cite{meier} to be able to
determine the entire architecture. Further, as shown in Figure \ref{convergence_plot}
that the loss temporarily suffers from the removal of neurons. It is likely
that the loss would be more stable if the number of neurons converged faster or
neurons disappeared more slowly. For this reason we plan to explore proximal
gradient methods to optimize the filter vectors and/or randomize neuron
removals. Finally, during our evaluation we picked small datasets mainly to be able
to train many models and have statistically significant distributions. We 
plan to verify that our approach
 see if it generalizes to bigger
datasets and other architectures like \texttt{ResNet} ~\cite{He2016}, which is possible
with small
modifications to our existing code base.
